SIMD\slash SIMT-vectorized Sparse Chemical Kinetic Jacobian\slash Thermo-chemical Source Term Evaluation

### 2018-02-23 paper.tex
add initial finite difference jacobian stuff

The mean runtime per-condition for the sparse and dense Jacobian evaluation on the \gpunew/ GPU is plotted in~\cref{F:gpu_sparse_vs_dense}.
As in the species evaluation case, the mean runtime per-condition drops steadily, for both cases becoming constant after just \num{3e3} states (except for \ce{H2}\slash\ce{CO}, which levels off near \num{e4} conditions).
\Cref{F:gpu_jacobian_speedup} presents the speedup of the \gpunew/ over the \gpuold/ GPU for sparse\slash dense Jacobian evaluation; sparse evaluation on the \gpunew/ is \SIrange{1.22}{1.63}{$\times$} faster, while dense evaluation is \SIrange{1.62}{2.1}{$\times$} faster.
The speedup on the \gpunew/ appears to decrease with increasing mechanism size for the sparse evalution, however we note that the \gpuold/ GPU has high variability in runtimes for the larger models.\todo{missing srv2 data for ipentanol, and need to rerun srv2 data}.

@@

>>>>

The mean runtime per-condition for the sparse and dense Jacobian evaluation on the \gpunew/ GPU is plotted in~\cref{F:gpu_sparse_vs_dense}.
As in the species evaluation case, the mean runtime per-condition drops steadily, for both cases becoming constant after just \num{3e3} states (except for \ce{H2}\slash\ce{CO}, which levels off near \num{e4} conditions).
\Cref{F:gpu_jacobian_speedup} presents the speedup of the \gpunew/ over the \gpuold/ GPU for sparse\slash dense Jacobian evaluation; sparse evaluation on the \gpunew/ is \SIrange{1.22}{1.63}{$\times$} faster, while dense evaluation is \SIrange{1.62}{2.1}{$\times$} faster.
The speedup on the \gpunew/ appears to decrease with increasing mechanism size for the sparse evalution, however we note that the \gpuold/ GPU has high variability in runtimes for the larger models.\todo{missing srv2 data for ipentanol, and need to rerun srv2 data}

@@

In~\cref{F:fd_vs_analytical}, the performance of the sparse analytical-Jacobian is compared to a sparse first-order finite-difference Jacobian on both the \avx/ CPU and the \gpuold/ \gpunew/ GPUs.\todo{missing data}
\Cref{F:fd_vs_analytical_cpu} shows large speedups for both the OpenMP code and shallow-vectorized OpenCL code; the analytical OpenMP Jacobian is \SIrange{3.92}{-11111}{$\times$} faster, while the analytical OpenCL Jacobian achives speedups of \SIrange{17}{55}{$\times$}.
It is noted that the current finite-difference formulation breaks Intel's auto-vectorizer, hence the OpenCL speedup comparison is made against the unvectorized OpenCL code; it is not a high priority to fix this issue, as the finite-difference Jacobian is implemented for comparison purposes only.
Although we do not display the dense finite-difference Jacobian speedup in~\cref{F:fd_vs_analytical_cpu}, the dense OpenCL and OpenMP analytical Jacobian codes out-perform the finite-difference variants by even larger margins; \SIrange{24}{-155}{$\times$} for OpenCL, and \SIrange{10}{-11111}{$\times$} for OpenMP.
A comparison of the sparse analytical and finite-difference Jacobians is presented in~\cref{F:fd_vs_analytical_gpu}.
The analytical Jacobian on the \gpunew/ shows large speedups, increasing with chemical model size, in the range of \SIrange{5}{16}{$\times$}; however, the \gpuold/ GPU shows almost no speedup, it is not clear why this is the case.\todo{reaaaally need to re-run this data}
Although not pictured, the dense analytical-Jacobian on the \gpunew/ GPU has larger speedups compared with the dense finite-difference Jacobian, in the range of \SIrange{5}{42}{$\times$}.

\begin{figure}[htbp]
   \centering
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{finite_difference_vs_analytical.pdf}
      \caption{Mean runtime per-condition of sparse and dense Jacobian evaluation on the \gpunew/ GPU.}
      \label{F:fd_vs_analytical_cpu}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.48\linewidth}
      \includegraphics[width=\textwidth]{finite_difference_vs_analytical_gpu.pdf}
      \caption{Speedup of the \gpunew/ over \gpuold/ GPUs for sparse and dence Jacobian evaluation.}
      \label{F:fd_vs_analytical_gpu}
  \end{subfigure}
  \caption{The performance of a sparse, first-order forward finite-difference Jacobian compared to the analytical Jacobian on the \avx/ CPU and both GPUs.}
  \label{F:fd_vs_analytical}
\end{figure}
